---
title: "607 Week 9 Assignment Web APIs"
author: "Brandon Chung"
date: "`r Sys.Date()`"
output: html_document
---

## Instructions


The New York Times web site provides a rich set of APIs, as described here:  https://developer.nytimes.com/apis 
You’ll need to start by signing up for an API key. 
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and 
transform it into an R DataFrame.


```{r}
#Load packages
library(httr2)
library(jsonlite)
library(tidyverse)
library(infer)
```

```{r}
# The API call from NYTs requires API key to be put into the url at the end, and certain parameters for querying eg (lists/names)
# API call to find the list of best sellers that I want to research

api_key <- "HkhZiJEhn1oSfOE6SGZVqKWDyvYsCy7r"

url <- "https://api.nytimes.com/svc/books/v3/lists/names.json?api-key=HkhZiJEhn1oSfOE6SGZVqKWDyvYsCy7r"

#Using json to perform the api call

json_data <- fromJSON(url)

# Flatten the json data

flat_json <- flatten(json_data)
                     
#Reading json data response into a data frame in R

df <- as.data.frame(flat_json$list_name_encoded)

head(df)
```




```{r}
url2 <- "https://api.nytimes.com/svc/books/v3/lists/current/business-books.json?api-key=HkhZiJEhn1oSfOE6SGZVqKWDyvYsCy7r"


json_data2 <- fromJSON(url2)

# Flatten the json data

                     
#Reading json data response into a data frame in R

df2 <- as.data.frame(json_data2$results$books)

head(df2)
```

## Analysis

I want to test if the top 5 books have a larger image size than the bottom 5 from the top 10 best selling buisness books.

```{r}
df2 <- df2[,c(1,11,16,17)]

df2 <- df2 %>%
  mutate(image_area = book_image_width * book_image_height) %>%
  mutate(top.5 = rank < 6)
```

```{r}
ggplot(df2, aes(x = top.5, y = image_area)) +
  geom_boxplot()
```
