---
title: "607 Week 11 Extra Credit Recommender Systems"
author: "Brandon Chung"
date: "`r Sys.Date()`"
output: html_document
---

## Instructions
Using the information you collected on movie ratings, implement a Global Baseline Estimate recommendation system in R.  

Most recommender systems use personalized algorithms like “content management” and “item-item collaborative filtering.” Sometimes non-personalized recommenders are also useful or necessary. One of the best non-personalized recommender system algorithms is the “Global Baseline Estimate.

The job here is to use the survey data collected and write the R code that makes a movie recommendation using the Global Baseline Estimate algorithm.  Please see the attached spreadsheet for implementation details.

## Introduction

I will create a recommender system based on the given movie ratings data and Global Baseline Estimate algorithm. I will show what movie each critic would most likely enjoy out of the movies that they have not seen from the 6 movies in the data.

The global baseline estimate algorithm is GBE = Mean Movie Rating + SPecific movie rating relative to average + specific critic's average relative to the mean movie rating.

```{r}
library(janitor)
library(tidyverse)
```
```{r}
# Importing dataset

movieratings <- read.csv("https://raw.githubusercontent.com/Chung-Brandon/607/refs/heads/main/movieratings.csv")

# Removing empty rows
movieratings <- movieratings[c(1:16),]

# Cleaning column names to be undercase
movieratings <- movieratings %>%
  clean_names()
```

```{r}
# Viewing the movie rating matrix
head(movieratings)
```

```{r}
# Tidying the data
movieratings.tidy <- movieratings %>%
  pivot_longer(
    cols = 2:7,
    names_to = "movie",
    values_to = "rating"
  )

head(movieratings.tidy)
```

```{r}
# creating df for critic average ratings 

critic.avg <- movieratings.tidy %>%
  group_by(critic) %>%
  summarize(avg.rating = mean(rating, na.rm = TRUE))

head(critic.avg)
```

```{r}
# Calculating the mean_movie rating
mean_movie_rating <- mean(movieratings.tidy$rating, na.rm = TRUE)

mean_movie_rating
```
```{r}
# Mutate critic rating relative to mean_movie_rating into critic.avg

critic.avg <- critic.avg %>%
  mutate(critic.mean_relativerating = avg.rating - mean_movie_rating)
```
```{r}
# Calculate movie rating relative to mean_movie_rating

movie.avg <- movieratings.tidy %>%
  group_by(movie) %>%
  summarise(avg.rating = mean(rating, na.rm = TRUE))

movie.avg <- movie.avg %>%
  mutate(movie.mean_relativerating = avg.rating - mean_movie_rating)
```
```{r}
# Calculating global baseline estimate for each movie

# Join the datasets together for relevent values for GBE calculation
combined <- movieratings.tidy %>%
  left_join(critic.avg, by = "critic") %>%
  left_join(movie.avg, by = "movie")


# Drop unnecessary columns
combined <- combined[,-c(4,6)]

# Mutate in GBE column
combined <- combined %>%
  mutate(gbe = mean_movie_rating + critic.mean_relativerating + movie.mean_relativerating)

head(combined)
```
```{r}
# Creating a recommendations list where rating = NA, and arranging by GBE

recommendations <- combined %>%
  filter(is.na(rating)) %>%
  group_by(critic) %>%
  arrange(desc(gbe)) %>%
  select(critic, movie, gbe)
  

top.recommendation <- combined %>%
  filter(is.na(rating)) %>%
  group_by(critic) %>%
  slice_max(gbe, n = 1) %>%
  select(critic, movie, gbe)

head(top.recommendation)
```
```{r}
top.recommendation %>%
  group_by(movie) %>%
  count() %>%
  arrange(desc(n))
```


## Conclusion 

In conclusion, using the Global Baseline Estimate recommendation system I have found the top recommended movies for each critic, if they have not rated any of the 6 movies listed, and displayed them in the data set top.recommendation. Deadpool rated the highest recommended movie for 7 out of 12 critics, and jungle_book came in second being top rated for 2 out of 12 critics.  
