---
title: '607 Final Project: Stock Price Movement vs Earnings'
author: "Brandon Chung"
date: "12/18/2024"
output:
  slidy_presentation: default
  beamer_presentation: default
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Loading packages

library(tidyverse)
library(rvest)
library(stringr)
library(kableExtra)
library(lubridate)
library(janitor)
```


## Overview

I acquired, transformed and analyzed historical stock price data and reported earnings per share (EPS) for the companies Apple (AAPL), Applovin (APP), and Nividia (NVDA). I performed a hypothesis test on the average percent price change per quarter. 

Hypothesis: The quarter following a reported +30% EPS surprise has a greater price appreciation than if the EPS surprise is <30%

\[ H_0: \mu = \mu_0 \]

\[ H_a: \mu > \mu_0 \]

## Project Motivation

My project motivation was to deepen my understanding of stock price action to better inform my own investment choices, but also to practice my data science skills on a data set I find challenging.  

## Data Acquistion - Price History

Stock price data is available for download at nasdaq.com

```{r, out.width='80%', fig.align='center', echo=TRUE}
# Reading in price history raw data csv from github
appl_ph <- read.csv("https://raw.githubusercontent.com/Chung-Brandon/607/refs/heads/main/AAPL_ph.csv")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
head(appl_ph) %>%
  kable("latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped","scale_down"))

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Reading in price history for NVDA and APP stocks
nvda_ph <- read.csv("https://raw.githubusercontent.com/Chung-Brandon/607/refs/heads/main/NVDA_ph.csv")

app_ph <- read.csv("https://raw.githubusercontent.com/Chung-Brandon/607/refs/heads/main/APP_ph.csv")
```


## Data Acquistion - Reported Earnings per Share
Data scraping with rvest on alphaquery.com

```{r, out.width='60%', fig.align='center', echo=TRUE}
url <- "https://www.alphaquery.com/stock/NVDA/earnings-history"
webpage <- read_html(url)

earnings_table <- html_nodes(webpage, 'table')
nvda.eps <- html_table(earnings_table)[[1]]
```

## Data Acquistion - Reported Earnings per Share

```{r echo=FALSE, message=FALSE, warning=FALSE}
head(nvda.eps) %>%
  kable("latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
url <- "https://www.alphaquery.com/stock/APP/earnings-history"
webpage <- read_html(url)

earnings_table <- html_nodes(webpage, 'table')
app.eps <- html_table(earnings_table)[[1]]

url <- "https://www.alphaquery.com/stock/AAPL/earnings-history"
webpage <- read_html(url)

earnings_table <- html_nodes(webpage, 'table')
appl.eps <- html_table(earnings_table)[[1]]
```

## Data Transformation - Clean and Mutate PH

```{r}
# Mutate in the company ticker
app_ph <- app_ph %>%
  mutate(company = "app") %>%
  relocate(company, .before = 1)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
appl_ph <- appl_ph %>%
  mutate(company = "appl") %>%
  relocate(company, .before = 1)

nvda_ph <- nvda_ph %>%
  mutate(company = "nvda") %>%
  relocate(company, .before = 1)
```
```{r}
# Bind price history tables together
combined_ph <- rbind(app_ph, appl_ph, nvda_ph)
```

```{r}
# Removing unnecessary columns
combined_ph <- combined_ph[,(1:3)] %>%
  rename(close = Close.Last,
         date = Date)
```
## Data Transformation - Clean and Mutate PH

```{r echo=FALSE, message=FALSE, warning=FALSE}
head(combined_ph) %>%
  kable("latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped","scale_down"))
```

## Data Transformation - Clean and Mutate EPS

```{r echo=FALSE, message=FALSE, warning=FALSE}
app.eps <- app.eps %>%
  mutate(company = "app") %>%
  relocate(company, .before = 1)

appl.eps <- appl.eps %>%
  mutate(company = "appl") %>%
  relocate(company, .before = 1)

nvda.eps <- nvda.eps %>%
  mutate(company = "nvda") %>%
  relocate(company, .before = 1)

```
```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_eps <- rbind(app.eps,appl.eps,nvda.eps) %>%
  clean_names()
```

```{r, size='small'}
# Converting EPS data into numeric data for calculations
combined_eps <- combined_eps %>%
  mutate(estimated_eps = as.numeric(gsub("[^0-9.]", "", estimated_eps)),
         actual_eps = as.numeric(gsub("[^0-9.]", "", actual_eps)))
```

```{r, size='tiny'}
# Mutating in EPS surpise %
combined_eps <- combined_eps %>%
  mutate(eps.surprise.p = (((actual_eps - estimated_eps) / estimated_eps) * 100))
```

```{r, size='tiny'}
# Removing unneccessary columns
combined_eps <- combined_eps[,-c(3,4,5)]
```

## Data Transformation - Clean and Mutate EPS

```{r echo=FALSE, message=FALSE, warning=FALSE}
head(combined_eps) %>%
  kable("latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped","scale_down"))

```


## Data Transformation - Merging and Forming Tidy Table

```{r, size='tiny'}
# standardizing date columns in both combined data frames
combined_eps <- combined_eps %>%
  rename(date = announcement_date)

combined_eps$date <- ymd(combined_eps$date)
combined_ph$date <- mdy(combined_ph$date)

```
```{r, size='tiny'}
## Using innerjoin to keep only the observations with dates of EPS announcement
tidy.df <- combined_ph %>%
  inner_join(combined_eps, by = c("company", "date"))
```

## Tidy Data

```{r, size='tiny'}
# Converting close variable to numeric data
tidy.df <- tidy.df %>%
  mutate(close = as.numeric(gsub("[^0-9.]", "", close)))
```

```{r, size='tiny'}
# Calculating the percentage change from EPS announcement to next EPS announcement
tidy.df <- tidy.df %>%
  group_by(company) %>%
  arrange(company, date) %>%
  mutate( percentage_change = (lead(close) - close) / close * 100 )
```

## Tidy Data 

Left with 95 observations of 5 variables

```{r echo=FALSE, message=FALSE, warning=FALSE}
head(tidy.df) %>%
  kable("latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
tidy.df %>%
  group_by(company) %>%
  summarize(count = n())
```

## Data Visualization

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(tidy.df, aes(x = date)) +
  geom_line(aes(y = percentage_change, color = "Percentage Change")) +
  geom_line(aes(y = eps.surprise.p, color = "EPS Surprise")) +
  labs(title = "Two-Line Plot Faceted by Company", x = "Date", y = "Value") +
  scale_color_manual(values = c("Percentage Change" = "blue", "EPS Surprise" = "red")) + facet_wrap(~ company) + theme_minimal()
```

## Data Visualization

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(na.omit(tidy.df), aes(x = eps.surprise.p, y = percentage_change)) +
  geom_point()
```


## Analysis

\[ H_0: \] The percentage change for the period after a >30% EPS surprise is the same for <30% EPS surprise.
\[ H_a: \] The percentage change for the period after a >30% EPS surprise is greater than for a <30% EPS surprise.


\( \mu: \) 14.00

\( \mu_{30}: \) 51.467


## Analysis 

```{r}
tidy.df <- tidy.df %>%
  mutate(eps.30 = ifelse(percentage_change > 30, "Yes", "No"))
tidy.df %>%
  group_by(eps.30) %>%
  summarise(mean_pchange = mean(percentage_change),
            count = n())
```

## Analysis

```{r}
mean_p.change <- mean(tidy.df$percentage_change, na.rm = TRUE)
mean_p.change
```


## Analysis 

Calculating a 95% confidence interval around the eps.30 "Yes" group

```{r}
# Creating a CI with the t.test function

t_test <- tidy.df[,5]
t_test_result <- t.test(t_test, conf.level = 0.95)
conf_interval <- t_test_result$conf.int
cat("95% Confidence Interval: [", conf_interval[1], ", ", conf_interval[2], "]\n")
```

## Analysis visualization

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(na.omit(tidy.df), aes(eps.30, percentage_change)) +
  geom_boxplot()
```


## Conclusion

Because our EPS surprise >30% group has a mean percent change of 51, outside of the calculated CI, we reject the null hypothesis and can say that we are 95% confident that for the quarter post EPS announcement, if there is an EPS surprise of 30% or greater, we will likely see greater percent stock appreciation.
