---
title: "607 Week 4 Data Structures in R"
author: "Brandon Chung"
date: "`r Sys.Date()`"
output: html_document
---

Introduction:

This week's assignment is about tidying and transforming data in R using the tidyr and dplyr packages. I will attempt to import the given data in its original form into MySQL Workbench and tidy the data in R. 

```{r}
#reading the csv file that I created into R

library("RMySQL")

mydb <- dbConnect(MySQL(),user= 'bchung',password= 'august29!',dbname= '607',host= 'bchungcuny607.mysql.database.azure.com')

dbListTables(mydb)

#reading the data table into R
Data <- dbReadTable(mydb,"tidyingdata")

#the data was imported with the unnamed variables being assigned MyUnknownColumn and MyUnknownColumn_.0. by R studio.
```


```{r}
library(dplyr)
library(tidyr)

# Removing the third (blank row) observation in the data frame
Data <- Data[-c(3),]
```


```{r}
# Renaming the originally blank column names
Data <- Data %>%
  rename(airline = MyUnknownColumn,
         status = MyUnknownColumn_.0.)
```
```{r}
# Mutating variable Airline to fill in the blank values
Data <- Data %>%
  mutate(airline = c('ALASKA','ALASKA','AM_WEST','AM_WEST'))
```
```{r}
# Pivoting long for analysis
Data <- Data %>% 
  pivot_longer(
    cols =  3:7,
    names_to = 'destination',
    values_to = 'count')
```
```{r}
# Relocating variable
Data <- Data %>%
  relocate(status, .after = destination)

# Pivoting wide 
Data <- Data %>% 
  pivot_wider(
    names_from = status,
    values_from = count)
```
```{r}
# Cleaning variable names
colnames(Data) <- gsub(" ", "_", colnames(Data))
```


```{r}
# Converting character data types into number for analysis
Data$on_time <- as.numeric(gsub(",", "", Data$on_time))
Data$delayed <- as.numeric(as.character(Data$delayed))
```

```{r}
# Calculating the delay ratios for each airline / destination combination
Data <- Data %>% 
  mutate(delayratio = delayed/on_time)
Data <- Data %>%
  mutate(total_flights = on_time + delayed)
Data
```

```{r}
# The code below shows the difference in mean and median delay ratios across the two airlines for all destination. Alaska Airlines shows less of a chance for their flights to be delayed compared to AM_West.

Data %>%
  group_by(airline) %>%
  summarise(mean_d.ratio = mean(delayratio),
            median_d.ratio = median(delayratio))
```

```{r}
#PLotting the delay ratios of Alaska Airlines by destination
library(ggplot2)
ggplot(subset(Data, airline %in% c("ALASKA")), 
       aes(x = destination, y = delayratio)) +
  geom_point()
```
```{r}
#Plotting the delay ratios of AM West Airlines by destination
ggplot(subset(Data,airline %in% c("AM_WEST")),
       aes(x = destination, y = delayratio)) +
  geom_point()
```


Doing a quick eye comparison of the two dot plots above we can see that across both airlines the arrival delay ratios are aligned in terms of which desinations have more delays, however AM West shows higher delay ratios for each of the destinations.

## Conclusions:
Delays to these five destinations are likely impacted by non-airline related factors, as some destinations are delayed more than others regardless of the airline. Although, in comparison we can see that independent of the destination factor AM West shows a higher ratio of delayed to on time arrivals. If one wanted to avoid arrival delays I would recommend flying Alaska Airlines instead of AM West. 


