---
title: "607 Week 4 Data Structures in R"
author: "Brandon Chung"
date: "`r Sys.Date()`"
output: html_document
---

#Resubmission attempt for week 4's Data Structures in R assignment

Introduction:

This week's assignment is about tidying and transforming data in R using the tidyr and dplyr packages. I will attempt to import the given data in its original form into MySQL Workbench and tidy the data in R. 

```{r}
#reading the csv file that I created into R

library("RMySQL")

mydb <- dbConnect(MySQL(),user= 'bchung',password= 'august29!',dbname= '607',host= 'bchungcuny607.mysql.database.azure.com')

dbListTables(mydb)

#reading the data table into R
Data <- dbReadTable(mydb,"tidyingdata")

#the data was imported with the unnamed variables being assigned MyUnknownColumn and MyUnknownColumn_.0. by R studio.

Data
```


```{r}
library(dplyr)
library(tidyr)

# Removing the third (blank row) observation in the data frame
Data <- Data[-c(3),]
```


```{r}
# Renaming the originally blank column names
Data <- Data %>%
  rename(airline = MyUnknownColumn,
         status = MyUnknownColumn_.0.)
```
```{r}
# Mutating variable Airline to fill in the blank values
Data <- Data %>%
  mutate(airline = c('ALASKA','ALASKA','AM_WEST','AM_WEST'))

# Showing data frame after basic cleaning
Data
```
```{r}
# Pivoting long for analysis
Data <- Data %>% 
  pivot_longer(
    cols =  3:7,
    names_to = 'destination',
    values_to = 'frequency')
```

```{r}
# Converting character data types into number for analysis
Data$frequency <- as.numeric(gsub(",", "", Data$frequency))

```
# The tidy dataframe shows the observation of frequency (count of flights) per row. 
```{r}
Data
```

# Analysis
```{r}
# Calculating the delay ratios for each airline across all destinations
delayratio.airline <- Data %>%
  group_by(airline,status) %>%
  summarise(frequency = sum(frequency)) %>%
  mutate(delay.ratio = frequency / sum(frequency)) %>%
  filter(status == "delayed")

delayratio.airline
```

As we can see, Alaska Airlines has a higher ratio of delayed flights when compared to AM_West even with a lower frequency of flights.


```{r}
#Comparing the delay ratios of each airline by destination can give us deeper insight to potential reasons for delays - if delays are typically caused by destination related factors or airline related factors.

delayratio.compare <- Data %>%
  group_by(airline, destination, status) %>%
  summarise(frequency = sum(frequency)) %>%
  mutate(delay.ratio = frequency / sum(frequency)) %>%
  filter(status =="delayed")
    
delayratio.compare  
```


```{r}
#PLotting the delay ratios by destination
library(ggplot2)
ggplot(delayratio.compare, aes(x = airline, y = delay.ratio)) +
  geom_point() +
  facet_wrap(vars(destination))
```

Doing a quick eye comparison of the graphs of delay ratios by destination above we can see that across both airlines the arrival delay ratios are aligned in terms of which desinations have more delays, however AM West shows higher delay ratios for each of the destinations. AM West possibly has airline related factors increasing delay ratios for San Francisco, Seattle, and San Diego.

## Conclusions:
Delays to these five destinations are likely impacted by non-airline related factors, as some destinations are delayed more than others regardless of the airline. Although, in comparison we can see that independent of the destination factor AM West shows a higher ratio of delayed to on time arrivals. If one wanted to avoid arrival delays I would recommend flying Alaska Airlines instead of AM West. 

